% =====================================================
% Literature Review
% =====================================================
\section{Literature Review}
% TODO: Provide an overarching narrative of related work in DR grading.

\subsection{Model Selection}
DR classification has largely tracked the broader ImageNet backbone progression. Gulshan \textit{et al.}~\cite{gulshan2016} set the template with an \textbf{Inception-v3}-based binary referable-vs-non-referable classifier that reached ophthalmologist-level sensitivity on 128k images, after which work on five-class severity grading shifted to deeper residual networks (\textbf{ResNet-50}, \textbf{DenseNet-121}) and compound-scaled \textbf{EfficientNet} variants. Across these systems the classification head is uniformly a flat softmax over Grades $0$--$4$, which discards the ordinal structure and penalises a Grade-$4$-as-Grade-$0$ confusion no more than a Grade-$4$-as-Grade-$3$ one---an assumption consistent with the high inter-grader disagreement reported on the same images by human ophthalmologists~\cite{krause2018}.

For our setting---a $\sim$3.6k-image training set on a single 4\,GB consumer GPU, with ordinality to be supplied by the loss rather than the backbone---the relevant axis is not raw capacity but the smallest backbone that still gives useful ImageNet-pretrained features without overfitting. \textbf{ResNet-50} ($\sim$25M parameters) brings roughly $2.2\times$ the parameters of ResNet-18 and, at this data scale, simply enlarges the overfitting surface; \textbf{EfficientNet-B0} ($\sim$5M parameters) is parameter-efficient on paper but its compound-scaling advantage requires more data than we have to be realised. \textbf{ResNet-18} ($\sim$11.2M parameters) sits at the sweet spot---deep enough to learn the multi-scale lesion textures that distinguish DR severity grades, shallow enough to fine-tune end-to-end in memory, and well-matched to the data regime so that gains from regularisation and ordinal-aware losses are not masked by capacity-driven overfitting. We therefore adopt \textbf{ResNet-18} as the backbone and direct the remaining effort towards the classification head, the loss function, and the preprocessing pipeline.


\subsection{Preprocessing}
Preprocessing is not a cosmetic step in medical imaging, it is the single largest lever between a model that learns lesion-level pathology and one that latches onto background artefacts. Raw fundus photographs vary enormously in illumination, contrast, white-balance, the size of the visible retinal disc, and the amount of black border surrounding it; without standardisation, the network is forced to spend representational capacity on these nuisance factors instead of on disease cues~\cite{salehi2023preprocessing,litjens2017survey}. 
A consistent finding in the medical CNN literature is that, in the absence of preprocessing, gradient-based attribution maps drift onto clinically irrelevant regions (background, large vessels, image borders), whereas appropriate preprocessing concentrates attention on the diagnostic structures themselves~\cite{salehi2023preprocessing}.

Within our pipeline, preprocessing is built around four mutually reinforcing components. \textbf{Geometric normalisation} resizes every image to a fixed $224\times224$ resolution and crops the circular field of view, removing the black corners that otherwise contribute uninformative gradients. \textbf{Photometric normalisation} subtracts the per-channel ImageNet mean $(0.485,0.456,0.406)$ and divides by its standard deviation $(0.229,0.224,0.225)$, ensuring that the feature statistics fed to the pretrained ResNet match those it was trained on, without this step, transfer learning is markedly less effective. \textbf{Data augmentation} (horizontal/vertical flips, small rotations, brightness-contrast jitter, CLAHE contrast enhancement) synthetically expands the effective training distribution and is one of the most powerful regularisers known for limited medical datasets~\cite{shorten2019aug,perez2017aug}. 
\textbf{Regularisation hyperparameters} that complement preprocessing, most notably \emph{dropout}, which randomly disables units during training to prevent co-adaptation~\cite{srivastava2014dropout}, and \emph{weight decay}, which penalises large parameter norms, directly address the overfitting we observed in the baseline. Closely related are the choice of \emph{weight initialisation} ~\cite{he2015delving}, the optimiser learning rate, and the use of pretrained ImageNet weights.

\subsection{Loss Function Definition}
% TODO: Compare cross-entropy, weighted CE, focal loss, ordinal regression losses,
% and CORAL framework.
Initially, the most obvious choice for a loss function is the standard \textbf{categorical cross-entropy}, which treats the five DR grades as independent classes. However, this approach discards the ordinal structure of the problem and fails to account for class imbalance, later introduced was \textbf{weighted cross-entropy} in~\cite{buda2018}, which assigns higher loss weights to the minority classes (Grades 3 and 4) to mitigate imbalance. \textbf{Focal loss}~\cite{lin2017focal} further refines this by down-weighting easy examples and focusing the model's attention on hard, misclassified samples, which is particularly beneficial in imbalanced settings. This brings our class imbalancement down to a manageable level, but it still treats the classes as independent.
\textbf{Ordinal regression losses} such as CORAL~\cite{cao2019coral} explicitly model the rank order of the classes by decomposing the problem into a series of binary classification tasks, ensuring that the model learns to respect the natural ordering of DR severity. This handles the severity of misclassification, for example, misclassifying Grade 4 as Grade 0 is penalised more than misclassifying it as Grade 3. This is a critical improvement over standard cross-entropy, which treats all misclassifications as equally bad.

\subsection{Evaluation Metric}
Plain accuracy is poorly suited to a five-class, severely imbalanced, ordinal task: predicting Grade~$0$ for every image already attains close to $50\%$ accuracy on APTOS-2019 without learning anything clinically useful. We therefore adopt the \textbf{Quadratic Weighted Kappa (QWK)} as the primary metric, following Cohen's weighted-kappa framework~\cite{cohen1968kappa}, which rewards agreement between predicted and true grades and penalises disagreement by the \emph{squared} distance between them, so a Grade-$4$-as-Grade-$0$ error is sixteen times costlier than an adjacent-grade error. QWK is the official scoring metric of both the Kaggle \emph{Diabetic Retinopathy Detection} (2015) and \emph{APTOS-2019 Blindness Detection} competitions, and de la Torre \textit{et al.}~\cite{delatorre2018kappa} have shown that it doubles as a differentiable training loss for DR grading. 

We supplement QWK with overall \textbf{accuracy} (for comparability with prior reports, while remembering its insensitivity to imbalance~\cite{buda2018}), \textbf{macro-F1} (equally-weighted per-class F1, which surfaces collapses on minority Grades~$3$ and~$4$), and the full \textbf{confusion matrix} (which reveals \emph{which} grades are being conflated and is what exposed the baseline's collapse onto Grades~$0$ and~$2$).
