% =====================================================
% References
% =====================================================
\begin{thebibliography}{99}

\bibitem{gulshan2016}
V. Gulshan \textit{et al.},
``Development and Validation of a Deep Learning Algorithm for Detection of Diabetic Retinopathy in Retinal Fundus Photographs,''
\textit{JAMA}, vol.~316, no.~22, pp.~2402--2410, 2016.

\bibitem{krause2018}
J. Krause \textit{et al.},
``Grader Variability and the Importance of Reference Standards for Evaluating Machine Learning Models for Diabetic Retinopathy,''
\textit{Ophthalmology}, vol.~125, no.~8, pp.~1264--1272, 2018.

\bibitem{fu2018dorn}
H. Fu, M. Gong, C. Wang, K. Batmanghelich, and D. Tao,
``Deep Ordinal Regression Network for Monocular Depth Estimation,''
in \textit{Proc. IEEE Conf. Computer Vision and Pattern Recognition (CVPR)}, 2018, pp.~2002--2011.

\bibitem{niu2016ordinal}
Z. Niu, M. Zhou, L. Wang, X. Gao, and G. Hua,
``Ordinal Regression with Multiple Output CNN for Age Estimation,''
in \textit{Proc. IEEE Conf. Computer Vision and Pattern Recognition (CVPR)}, 2016, pp.~4920--4928.

\bibitem{lin2017focal}
T.-Y. Lin, P. Goyal, R. Girshick, K. He, and P. Doll\'ar,
``Focal Loss for Dense Object Detection,''
in \textit{Proc. IEEE Int. Conf. Computer Vision (ICCV)}, 2017, pp.~2980--2988.
[Online]. Available: \url{https://arxiv.org/pdf/1708.02002v2}

\bibitem{buda2018}
M. Buda, A. Maki, and M. A. Mazurowski,
``A Systematic Study of the Class Imbalance Problem in Convolutional Neural Networks,''
\textit{Neural Networks}, vol.~106, pp.~249--259, 2018.

\bibitem{perez2017aug}
L. Perez and J. Wang,
``The Effectiveness of Data Augmentation in Image Classification using Deep Learning,''
\textit{arXiv preprint arXiv:1712.04621}, 2017.

\bibitem{selvaraju2017gradcam}
R. R. Selvaraju, M. Cogswell, A. Das, R. Vedantam, D. Parikh, and D. Batra,
``Grad-CAM: Visual Explanations from Deep Networks via Gradient-based Localization,''
in \textit{Proc. IEEE Int. Conf. Computer Vision (ICCV)}, 2017, pp.~618--626.

\bibitem{cao2019coral}
W. Cao, V. Mirjalili, and S. Raschka,
``Rank Consistent Ordinal Regression for Neural Networks with Application to Age Estimation,''
\textit{Pattern Recognition Letters}, vol.~140, pp.~325--331, 2020.
[Online]. Available: \url{https://sebastianraschka.com/pdf/papers/2019-coral1-2019-arxiv.pdf}

\bibitem{srivastava2014dropout}
N. Srivastava, G. Hinton, A. Krizhevsky, I. Sutskever, and R. Salakhutdinov,
``Dropout: A Simple Way to Prevent Neural Networks from Overfitting,''
\textit{Journal of Machine Learning Research}, vol.~15, no.~56, pp.~1929--1958, 2014.
[Online]. Available: \url{https://jmlr.org/papers/v15/srivastava14a.html}

\bibitem{he2015delving}
K. He, X. Zhang, S. Ren, and J. Sun,
``Delving Deep into Rectifiers: Surpassing Human-Level Performance on ImageNet Classification,''
in \textit{Proc. IEEE Int. Conf. Computer Vision (ICCV)}, 2015, pp.~1026--1034.

\bibitem{shorten2019aug}
C. Shorten and T. M. Khoshgoftaar,
``A Survey on Image Data Augmentation for Deep Learning,''
\textit{Journal of Big Data}, vol.~6, no.~60, pp.~1--48, 2019.
[Online]. Available: \url{https://link.springer.com/article/10.1186/s40537-019-0197-0}

\bibitem{salehi2023preprocessing}
A. W. Salehi \textit{et al.},
``Medical Image Analysis Using Deep Learning Algorithms,''
\textit{Frontiers in Public Health}, vol.~11, 2023.
[Online]. Available: \url{https://www.frontiersin.org/journals/public-health/articles/10.3389/fpubh.2023.1273253/full}

\bibitem{litjens2017survey}
G. Litjens \textit{et al.},
``A Survey on Deep Learning in Medical Image Analysis,''
\textit{Medical Image Analysis}, vol.~42, pp.~60--88, 2017.

\end{thebibliography}
